\subsection{device}\label{device}
The \verb|device| module provides a generic device abstraction for
byte-oriented input/output. A device is described by \verb|device_t| (Listing
\ref{device_t}) -- a table of function pointers for \verb|open|, \verb|read|,
\verb|write|, \verb|ioctrl|, \verb|close| and a \verb|ready_cb| callback,
complemented by application-specific \verb|user_data|. Concrete devices
register with \verb|device_init()| and are subsequently addressed via a
\verb|dev_handle_t| (valid when $\ge 0$); the \verb|device_*| functions check
the handle and delegate to the stored pointers. The module is used, among
others, by the \verb|serial| (Section \ref{serial}, \verb|serial_io|).

\begin{listing}
\begin{minted}{C}
typedef struct device_s {
    void *user_data;
    em_msg (*open)(dev_handle_t, void *user_data);
    em_msg (*read)(dev_handle_t, uint8_t *buffer, int16_t *cnt);
    em_msg (*write)(dev_handle_t, const uint8_t *buffer, int16_t cnt);
    em_msg (*ioctrl)(dev_handle_t, dev_command_t cmd, int16_t value);
    em_msg (*close)(dev_handle_t);
    em_msg (*ready_cb)(device_state_t state);
    uint8_t dev_type;
} device_t;
\end{minted}
\caption{Device interface \texttt{device\_t}}
\label{device_t}
\end{listing}

\noindent Which functions a device actually offers is expressed via the bit
field \verb|dev_func_t| (\verb|DEV_OPEN|, \verb|DEV_READ|, \verb|DEV_WRITE|,
\dots). \verb|device_check()| verifies that the function pointers required by
such a mask are set.

\subsubsection*{API}
{\sloppy\emergencystretch=3em
\begin{description}
\item[\texttt{device\_init(dev, user\_data)}] Registers the device and calls
  its \verb|open|. Returns: \verb|dev_handle_t| (the assigned handle, $0$ on
  error or a full table).
\item[\texttt{device\_check(hdl, dev\_type)}] Checks whether the functions
  required by the mask \verb|dev_type| are present. Returns: \verb|em_msg|
  (\verb|EM_OK| if all required pointers are set, otherwise \verb|EM_ERR|).
\item[\texttt{device\_read(hdl, buffer, cnt)} /
  \texttt{device\_write(hdl, buffer, cnt)}] Read or write via the device's
  function pointers. Returns: \verb|em_msg| (\verb|EM_ERR| if the device or the
  respective pointer is missing).
\item[\texttt{device\_reset(hdl)}] Resets the device to \verb|DEVICE_RESET|
  (all pointers \verb|NULL|). Returns: \verb|em_msg| (\verb|EM_OK|, or
  \verb|EM_ERR| for an unknown device).
\item[\texttt{device\_print(hdl)}] Prints the stored pointers for debugging.
  Returns: \verb|void|.
\end{description}
\par}
