\subsection{stateled}\label{stateled}
The \verb|stateled| module drives a row of status LEDs and thus makes the key
state visible. It maps the key states of a \verb|state_t| (Section \ref{state})
to LEDs: \verb|ON| is permanently lit, \verb|OFF| is off and \verb|BLINKING|
blinks -- the blink rate results from \verb|cycle_size|. In addition it
visualises the device's synchronisation state via the \verb|system_state_e|
(Section \ref{common}). The LEDs are driven through a \verb|gpio_port_t| (port
module); the constant \verb|REGUAL_STATELD| switches between normal and
inverted logic (active-low).

\begin{listing}
\begin{minted}{C}
typedef enum {
    led_0, led_1, led_2, led_3, led_4, led_5, led_6, led_7,
    led_fehler, led_normal,
} stateled_e;
\end{minted}
\caption{LED identifiers \texttt{stateled\_e}}
\label{stateled_e}
\end{listing}

\subsubsection*{API}
{\sloppy\emergencystretch=3em
\begin{description}
\item[\texttt{stateled\_init(state, port, cycle\_size, bli\_cnt)}]
  Initialises the module with the \verb|state_t| to display, a GPIO port (or
  the default assignment on \verb|NULL|), the cycle length for blinking and a
  blink counter; runs a running-light test across all LEDs as a check.
  Returns: \verb|void|.
\item[\texttt{stateled\_deinit()}] Deactivates the module. Returns: \verb|void|.
\item[\texttt{stateled\_on(nr)} / \texttt{stateled\_off(nr)} /
  \texttt{stateled\_all\_off()}] Switch a single LED or all LEDs. Returns:
  \verb|em_msg| (\verb|EM_ERR| if the module is not initialised, otherwise
  \verb|EM_OK|).
\item[\texttt{stateled\_toggle\_pin(nr)} / \texttt{stateled\_toggle\_port()}]
  Invert a single LED or all of them together. Returns: \verb|em_msg|
  (\verb|EM_ERR| if not initialised).
\item[\texttt{stateled\_set(val)} / \texttt{stateled\_set\_mask(mask)}] Set the
  LED pattern directly or define the active bit mask. Returns: \verb|em_msg|
  (\verb|EM_ERR| if not initialised).
\item[\texttt{stateled\_iterate()}] Advances cyclically to the next LED
  (running light). Returns: \verb|void|.
\item[\texttt{stateled\_show(state)}] In the \verb|SYNCHRONIZE_READY| or
  \verb|SYNCHRONIZE_DOING| state, displays the \verb|state_t| on the LEDs
  (on/off/blinking) and signals errors by toggling the entire port. Returns:
  \verb|void|.
\item[\texttt{stateled\_update(state)}] Variant that drives a running light or
  an error indication depending on the synchronisation state. Returns:
  \verb|bool|.
\end{description}
\par}
