\subsection{xpad}\label{xpad}
The \verb|xpad| module is a concrete \verb|kybd_t| implementation (Section
\ref{keyboard}) for keys connected via GPIO. It provides two devices:
\begin{description}
\item[\texttt{xscan\_dev}] a $4\times4$ matrix keypad that is scanned column by
  column (\verb|XSCAN|).
\item[\texttt{eight\_dev}] eight directly connected keys (\verb|EIGHTKEY|).
\end{description}
The columns (\verb|spalte|) are outputs, the rows (\verb|zeile|) are inputs
with pull-up; both are addressed via the \verb|gpio_port_t| (port module). Each
key passes through a debounce based on \verb|mkey_t| (Listing \ref{mkey_t}): a
level is only considered stable after \verb|STABLE_CNT| identical samples and
is then taken over into the device's \verb|state_t|.

\begin{listing}
\begin{minted}{C}
typedef struct key_s {
    bool last;      // last stable level
    bool current;   // currently read level
    bool unstable;  // level has changed
    uint8_t cnt;    // counter of identical samples
    bool stable;    // debounced level
} mkey_t;

typedef struct xpad_dev_s {
    gpio_port_t spalte;        // columns (outputs)
    gpio_port_t zeile;         // rows (inputs)
    dev_type_e dev_type;
    mkey_t key[MAX_BUTTON_CNT]; // debounce state per key
    state_t state;             // key state
} xpad_dev_t;
\end{minted}
\caption{Debounce and device structure \texttt{mkey\_t} / \texttt{xpad\_dev\_t}}
\label{mkey_t}
\end{listing}

\subsubsection*{Scanning}
{\sloppy\emergencystretch=3em
For the matrix (\verb|xscan_dev|) the output is activated per column, after
\verb|SETTLE_TIME_MS| settling time all rows are read and mapped via
\verb|(zeile, spalte)| to a key index. For \verb|eight_dev| each key is read
directly. Every read level passes through the debounce; once a key is detected
as stably pressed, the module propagates the change into the \verb|state_t| and
sets its \emph{dirty} flag. The assignment visible to the application is
defined by the labels in the \verb|state| (e.g.\ \verb|1 2 3 A / 4 5 6 B / ...|
for the matrix).
\par}

\noindent The actual vtable methods (\verb|xpad_init|,
\verb|xpad_spalten_scan|, \verb|xpad_eight_scan|, \verb|xpad_state|,
\verb|xpad_diff|, \verb|xpad_add|, \dots) are file-local and are reached
through the exported \verb|kybd_t| \verb|xscan_dev| and \verb|eight_dev| as
well as the \verb|keyboard_*| functions. Without a supplied configuration,
\verb|xpad_init()| uses a fixed default pin assignment (GPIOA/B/C pins).
